# Tests ########################################################################


# Load Packages
```{r}
library(tidyverse)
library(googlesheets4)
library(meta)
library(openssl)
library(gridExtra)
library(metafor)
library(robumeta)
library(knitr)
library(stargazer)
library(weightr)
library(MetaUtility)
```

# Religion (trt) should be _protective_ and, therefore, we should expect
# non-negative values

```{r Data}
d.converted<- subset(d, d$vi != 'NA')
d.converted<- d
d.converted <- d.converted %>% arrange(yi) # sort by point estimate
#d.converted <- d.converted %>% arrange(vi) # sort by variance

# Filter out problematic papers (for now)
# Mendolia et al. 2019
# Vogelsang and Lariscy 2020
d.converted <- d.converted %>% filter(authoryear != 'Mendolia et al. 2019')
d.converted <- d.converted %>% filter(authoryear != 'Vogelsang and Lariscy 2020')

```

# Create data frames (all, by outcome)
```{r}
# All
d.all <- d.converted
d.all <- d.all %>% filter( !(authoryear == 'Burdette et al. 2018' & substudy == 'Marijuana') )
d.all <- d.all %>% filter( !(authoryear == 'Hodge et al. 2021' & substudy == 'Marijuana') ) 
d.all <- d.all %>% filter( !(authoryear == 'Mak 2019' & substudy == 'Marijuana') ) 

# Alcohol
d.alcohol <- d.converted %>% filter(substudy == 'Alcohol')

# Smoking
d.smoking <- d.converted %>% filter(substudy == 'Smoking')

# Drugs
d.drugs <- d.converted %>% filter(substudy == 'Drugs')

# Marijuana
d.marijuana <- d.converted %>% filter(substudy == 'Marijuana')

```

# Descriptive Statistics
```{r}
# Outcome Counts - Extracted
outcome_counts_extracted <- df.substance %>%
  count(OutcomeSubType) %>%
  rename(Extracted = n)

# Outcome Counts - Used
outcome_counts_used <- d.converted %>%
  count(substudy) %>%
  rename(Used = n)

# Combine into a single table, matching by Outcome/Substudy
table2 <- full_join(outcome_counts_extracted, outcome_counts_used, 
                    by = c("OutcomeSubType" = "substudy")) %>%
  rename(Outcome = OutcomeSubType) %>%
  arrange(Outcome)

# Generate the Markdown table text for Table 2
markdown_table2 <- kable(
  table2, 
  format = "markdown", 
  align = c("l", "r", "r")
)

# Add explicit newlines for proper formatting
markdown_table2_text <- paste0(markdown_table2, collapse = "\n")

# Ensure newlines after each row
markdown_table2_text <- gsub("\\n", "\n", markdown_table2_text)

# Print Table 2 with proper newlines
cat(markdown_table2_text, "\n")
```

```{r}
# Missing Error Terms
missing_error_terms <- df.substance %>%
  summarise(True = sum(is.na(SE) & is.na(CIHigh)),
            False = sum(!is.na(SE) | !is.na(CIHigh))) %>%
  mutate(Category = "Missing Error Terms")

# Preventative vs Recovery
preventative_recovery <- df.substance %>%
  mutate(Preventative = grepl('prevent', PreventativeOrRecovery)) %>%
  count(Preventative) %>%
  tidyr::complete(Preventative = c(TRUE, FALSE), fill = list(n = 0)) %>%
  spread(Preventative, n) %>%
  rename(True = `TRUE`, False = `FALSE`) %>%
  mutate(Category = "Preventative")

# Headline in Abstract
headline_in_abstract <- df.substance %>%
  mutate(headline = grepl('yes', `Finding In Abstract?`)) %>%
  count(headline) %>%
  tidyr::complete(headline = c(TRUE, FALSE), fill = list(n = 0)) %>%
  spread(headline, n) %>%
  rename(True = `TRUE`, False = `FALSE`) %>%
  mutate(Category = "Headline")

# Combine into a single table
table1 <- bind_rows(missing_error_terms, preventative_recovery, headline_in_abstract) %>%
  select(Category, True, False)

# Generate the Markdown table text for Table 1
markdown_table1 <- kable(
  table1, 
  format = "markdown", 
  align = c("l", "r", "r")
)

# Add explicit newlines for proper formatting
markdown_table1_text <- paste0(markdown_table1, collapse = "\n")

# Ensure newlines after each row
markdown_table1_text <- gsub("\\n", "\n", markdown_table1_text)

# Print Table 1 with proper newlines
cat(markdown_table1_text, "\n")

```



# Basic Descriptive Values - Include Errors?
```{r Basic Descriptive Statistics}
# Missing Error Terms
df.substance %>% filter(is.na(SE) & is.na(CIHigh)) %>% count()
df.substance %>% filter(!is.na(SE) | !is.na(CIHigh)) %>% count()

# Preventative vs Recovery
df.substance$Preventative <- grepl('prevent', df.substance$PreventativeOrRecovery)
result_table <- df.substance %>%
 count( Preventative,  name = "Count")
print(result_table)

# Headline
df.substance$headline <- grepl('yes', df.substance$`Finding In Abstract?`)
result_table <- df.substance %>%
 count( headline,  name = "Count")
print(result_table)

# Outcome Counts - Extracted
result_table <- df.substance %>%
 count(OutcomeSubType, OutcomeSubType,  name = "Count") %>%  # Count unique tuples
  arrange(OutcomeSubType, OutcomeSubType, desc(Count))  # Sort the table
print(result_table)
# Outcome Counts - Used
result_table <- d.converted %>%
 count(substudy, substudy,  name = "Count") %>%  # Count unique tuples
  arrange(substudy, substudy, desc(Count))  # Sort the table
print(result_table)

```

# Forest Plots by Outcome
```{r Forest Plots by Outcome}
# Ensure the subgroup variable is a factor
d.converted$subgroup <- as.factor(d.converted$substudy)

# Get unique subgroups
unique_subgroups <- unique(d.converted$subgroup)

# Set up a 2x2 plotting area
par(mfrow = c(2, 2)) # 2 rows, 2 columns
par(mar = c(5, 5, 2, 5) + 0.1) # Adjusted margins for better spacing
par(cex = 0.8) # Increase text size for better readability

# Loop over subgroups and create a forest plot for each
for (grp in unique_subgroups) {
  subset_data <- d.converted[d.converted$subgroup == grp, ]
  
  # Extract effect sizes and variances for the subgroup
  effect_sizes <- subset_data$yi
  variances <- subset_data$vi
  study_labels <- subset_data$authoryear
  
  # Calculate confidence intervals
  ci.lb <- effect_sizes - 1.96 * sqrt(variances)
  ci.ub <- effect_sizes + 1.96 * sqrt(variances)
  
  # Create the forest plot for the subgroup with reversed vertical ordering
  forest(x = effect_sizes, vi = variances, slab = study_labels,
         xlab = "", 
         main = grp,
         alim = c(-4, 4), # Set x-axis limits to -4, 4
         at = seq(-4, 4, by = 1), # Set ticks at uniform intervals
         cex.lab = 1.2, # Increase the label size
         cex.axis = 0.5, # Increase the axis size
         rows = length(effect_sizes):1, # Reverse the order of the rows
         psize = 1.0, # Increase point size
         ilab.xpos = -4, # Position the labels closer to the plot
         xlim = c(-6, 6), # Extend xlim to give more space for labels
         digits = 2) # Limit the number of digits shown
}

# Reset the plotting area to single panel (if needed later)
par(mfrow = c(1, 1))
```


# Omnibus Random-effects model
```{r Fit random-effects model}
result <- rma(
  yi,  # Log-transformed RRs
  vi,  # Variances of the log-transformed RRs
  data = d.all,
  method = "REML"  # For random-effects model
)
summary(result) # Summary of the meta-analysis results


# Extract key statistics from the rma model
estimate <- formatC(result$b, format = "fg", digits = 3)
stderr <- formatC(result$se, format = "fg", digits = 3)
p_value <- ifelse(result$pval < 0.001, "<0.001", formatC(result$pval, format = "e", digits = 2))
ci_lower <- formatC(result$ci.lb, format = "fg", digits = 3)
ci_upper <- formatC(result$ci.ub, format = "fg", digits = 3)

# Model-level statistics
tau_sq <- formatC(result$tau2, format = "fg", digits = 3)
i2 <- formatC(result$I2, format = "fg", digits = 1)
num_outcomes <- result$k  # Number of studies
num_clusters <- "N/A"  # Not directly available in rma but could represent num_outcomes

# Create the inline text
inline_text <- paste0(
  "est = ", estimate, " (", ci_lower, ", ", ci_upper, "), ",
  "se = ", stderr, ", p = ", p_value, 
  ", I^2 = ", i2, ", \\tau^2 = ", tau_sq, 
  ", n = ", num_outcomes
)

# Print the inline text for markdown
cat(inline_text, "\n")

# Extract key statistics
estimate <- result$b

# Exponentiate the estimate
exp_estimate <- exp(estimate)

# Format the result for printing
formatted_estimate <- formatC(estimate, format = "fg", digits = 4)  # Rounded to 4 digits
formatted_exp_estimate <- formatC(exp_estimate, format = "fg", digits = 2)  # Rounded to 2 digits

# Create the standalone text for the exponentiated estimate
exp_text <- paste0("exp(", formatted_estimate, ") = ", formatted_exp_estimate)

# Print the standalone exponentiated estimate
cat(exp_text, "\n")

# Extract heterogeneity statistics
Q_stat <- formatC(result$QE, format = "fg", digits = 4)  # Q statistic
Q_df <- result$k - 1  # Degrees of freedom (df = number of studies - 1)
Q_pval <- ifelse(result$QEp < 0.001, "<.0001", formatC(result$QEp, format = "fg", digits = 4))  # p-value for Q

# Extract tau^2 and its SE
tau_sq <- formatC(result$tau2, format = "fg", digits = 4)  # tau^2
tau_sq_se <- formatC(result$se.tau2, format = "fg", digits = 4)  # SE of tau^2

# Print the heterogeneity stats for easy copying
heterogeneity_text <- paste0("Q(df = ", Q_df, ") = ", Q_stat, ", p-val = ", Q_pval)
tau_sq_text <- paste0("tau^2 = ", tau_sq, " (SE = ", tau_sq_se, ")")

# Print the heterogeneity results separately
cat(heterogeneity_text, "\n")
cat(tau_sq_text, "\n")

```


# Omnibu Robust Meta
```{r}
( meta.rob = robu( yi ~ 1, 
                   data = d.all, 
                   studynum = as.factor(authoryear),
                   var.eff.size = vi,
                   modelweights = "HIER",
                   small = F) )

# Print results to show inline in the article 
num_clusters <- meta.rob$N
num_outcomes <- meta.rob$M
omega_sq <- formatC(meta.rob$mod_info$omega.sq, format = "fg", digits = 2)
tau_sq <- formatC(meta.rob$mod_info$tau.sq, format = "fg", digits = 2)

# Extract regression table results
estimate <- formatC(meta.rob$reg_table$b.r, format = "fg", digits = 3)
stderr <- formatC(meta.rob$reg_table$SE, format = "fg", digits = 3)
p_value <- ifelse(meta.rob$reg_table$prob < 0.001, "<0.001", formatC(meta.rob$reg_table$prob, format = "e", digits = 2))
ci_lower <- formatC(meta.rob$reg_table$CI.L, format = "fg", digits = 3)
ci_upper <- formatC(meta.rob$reg_table$CI.U, format = "fg", digits = 3)

# Create the inline text
inline_text <- paste0(
  "est = ", estimate, " (", ci_lower, ", ", ci_upper, "), ",
  "se = ", stderr, ", p = ", p_value, 
  ", \\omega^2 = ", omega_sq, ", \\tau^2 = ", tau_sq, 
  ", n = ", num_outcomes, ", k = ", num_clusters
)

# Print the inline text
cat(inline_text, "\n")
```

# Main Figure: Make forest plot for omnibus meta robust results
```{r}
library(metafor)

# Sort the data in descending order by effect size
d.all <- d.all[order(d.all$yi, decreasing = TRUE), ]

# Extract necessary data for the forest plot
effect_sizes <- d.all$yi
variances <- d.all$vi
study_labels <- d.all$authoryear

# Calculate confidence intervals
ci.lb <- effect_sizes - 1.96 * sqrt(variances)
ci.ub <- effect_sizes + 1.96 * sqrt(variances)

# Extract overall mean effect size and its confidence interval
mean_effect <- meta.rob$reg_table$b.r
mean_ci.lb <- meta.rob$reg_table$CI.L
mean_ci.ub <- meta.rob$reg_table$CI.U

# Set the output to a PNG file with appropriate dimensions
png("omnibus-robu-meta-forest-final.png", width = 8.5, height = 11, units = "in", res = 300)  

# Adjust the margins and set smaller text sizes
par(mar = c(4, 4, 0, 2))  # Bottom, Left, Top, Right

# Create the forest plot with optimized spacing and smaller text sizes
forest(effect_sizes, vi = variances, slab = study_labels,
       xlab = "Effect Size", 
       alim = c(min(ci.lb, mean_ci.lb), max(ci.ub, mean_ci.ub)),
       xlim = c(min(ci.lb, mean_ci.lb) - 1.5, max(ci.ub, mean_ci.ub) + 1.5),  # Increased space around plot
       cex = 0.4,  # Smaller text size for study labels
       psize = 0.8,  # Smaller point size
       ilab = cbind(format(round(ci.lb, 2)), format(round(ci.ub, 2))),  # Add CI labels
       ilab.xpos = c(min(ci.lb, mean_ci.lb) - 2, max(ci.ub, mean_ci.ub) + 2),  # More space for CI labels
       ilab.pos = 2,  # Position of the labels relative to the plot
       header = "Study",
       ylim = c(0, length(effect_sizes) + 4))  # Adjust vertical space for better fit

# Add a polygon to represent the overall effect size and confidence interval
addpoly(mean_effect, ci.lb = mean_ci.lb, ci.ub = mean_ci.ub, 
        rows = 0, cex = 0.4, mlab = "Overall Mean Effect", col = "gray")

# Add the vertical line at the null effect (0)
abline(v = 0, lty = 2, col = "red")

# Close the PNG device to save the plot
dev.off()
```


# Omnibus Funnel Plots & Test
```{r}
# Extract effect sizes and standard errors directly from the data
effect_sizes <- d.converted$yi
standard_errors <- sqrt(d.converted$vi)

# Identify non-outlier data (SE <= 0.5)
non_outliers <- standard_errors <= 0.5

# Create a two-panel figure
png("two_panel_funnel_plot.png", width = 12, height = 6, units = "in", res = 300)
par(mfrow = c(1, 2))  # Set up a 1x2 plotting area

# Panel A: Funnel plot with all data
funnel(effect_sizes, sei = standard_errors, 
       ylab = "Standard Error", xlab = "Effect Size", 
       main = "",  # No title
       pch = 19,  # Solid circles
       col = "black",  # Black points
       cex = 0.7,  # Reduce point size
       cex.lab = 0.8,  # Reduce label size
       cex.axis = 0.7)  # Reduce axis text size

# Panel B: Funnel plot without outliers (SE > 0.5)
funnel(effect_sizes[non_outliers], sei = standard_errors[non_outliers], 
       ylab = "Standard Error", xlab = "Effect Size", 
       main = "",  # No title
       pch = 19,  # Solid circles
       col = "black",  # Black points
       cex = 0.7,  # Reduce point size
       cex.lab = 0.8,  # Reduce label size
       cex.axis = 0.7)  # Reduce axis text size

# Close the PNG device to save the plot
dev.off()
```

# Egger's Test
```{r}
rma_model <- rma(yi, vi, data = d.all, method = "REML")
d.all$sei <- sqrt(d.all$vi)
egger_test <- rma(yi, vi, mods = ~ sei, data = d.all, method = "REML")
# Extract relevant statistics
intercept_est <- round(egger_test$b[1], 3)
intercept_se <- round(egger_test$se[1], 3)
intercept_pval <- ifelse(egger_test$pval[1] < 0.001, "<0.001", round(egger_test$pval[1], 3))
t_value <- round(egger_test$zval[1], 2)
ci_lower <- round(egger_test$ci.lb[1], 3)
ci_upper <- round(egger_test$ci.ub[1], 3)

# Generate markdown text
egger_results <- paste0(
  "Egger's test results: The intercept was estimated to be ", intercept_est,
  " (SE = ", intercept_se, "), with a 95% CI of [", ci_lower, ", ", ci_upper, 
  "], t(", egger_test$k-1, ") = ", t_value, ", p = ", intercept_pval, "."
)
# Print the markdown
cat(egger_results)
cat(egger_results, file = pipe("pbcopy"))
```

# Investigate Variance
```{r}
# Perform Egger's test on original data set
(result <- rma(
  yi,  # Log-transformed RRs
  vi,  # Variances of the log-transformed RRs
  data = d.all,
  method = "REML"  # For random-effects model
))

# Perform Egger's test for funnel plot asymmetry
egger_test <- regtest(result, model = "lm")
print(egger_test)

# Extract relevant statistics
t_value <- round(egger_test$zval, 4)
df <- egger_test$df
p_value <- ifelse(egger_test$pval < 0.001, "<0.001", round(egger_test$pval, 4))

# Generate markdown text for the test result
egger_results <- paste0(
  "Test for Funnel Plot Asymmetry: t = ", t_value, 
  ", df = ", df, 
  ", p = ", p_value
)
# Print the markdown
cat(egger_results)
# Optionally, copy to clipboard using pbcopy (for macOS)
cat(egger_results, file = pipe("pbcopy"))

# Investigate variance
plot(d.all$vi, main = "Variances Across Studies", xlab = "Study Index", ylab = "Variance", log = "y")

# Identify the studies with the largest variances
max_variance <- max(d.all$vi)
outlier_studies <- d.all[d.all$vi == max_variance, ]

# Print details of the outlier studies
print(outlier_studies)

# Define a threshold for exclusion (e.g., exclude the top 1% of variances)
threshold <- quantile(d.all$vi, 0.99)
d.all_filtered <- d.all[d.all$vi < threshold, ]

# Create a logical index for outliers
outliers <- d.all$vi >= threshold
plot(d.all$vi, 
     main = "Variances Across Studies", 
     xlab = "Study Index", 
     ylab = "Variance", 
     log = "y", 
     col = ifelse(outliers, "red", "black"),  # Red for outliers, black for others
     pch = 19)  # Use solid circles for better visibility


# Perform Egger's test again on the filtered dataset
(result <- rma(
  yi,  # Log-transformed RRs
  vi,  # Variances of the log-transformed RRs
  data = d.all_filtered,
  method = "REML"  # For random-effects model
))

# Perform Egger's test for funnel plot asymmetry
egger_test_filtered <- regtest(result, model = "lm")
print(egger_test_filtered)

# Extract relevant statistics
t_value <- round(egger_test_filtered$zval, 4)
df <- egger_test_filtered$df
p_value <- ifelse(egger_test_filtered$pval < 0.001, "<0.001", round(egger_test_filtered$pval, 4))

# Generate markdown text for the test result
egger_test_filtered <- paste0(
  "Test for Funnel Plot Asymmetry: t = ", t_value, 
  ", df = ", df, 
  ", p = ", p_value
)

# Print the markdown
cat(egger_test_filtered)
# Optionally, copy to clipboard using pbcopy (for macOS)
cat(egger_test_filtered, file = pipe("pbcopy"))
```

# Meta Robu By Substudy
```{r}
d.all.byOutcome<- d.all %>% filter(subgroup != 'Combined')
d.all.byOutcome<- d.all.byOutcome %>% filter(subgroup != 'All Substances')
( meta.rob = robu( yi ~ 1 + substudy, 
                   data = d.all.byOutcome, 
                   studynum = as.factor(authoryear),
                   var.eff.size = vi,
                   modelweights = "HIER",
                   small = F) )
print(meta.rob)

# Extract the summary statistics for the model
extract_robu_results <- function(model) {
  # Extract the effect size, SE, CI, and p-value
  Effect_Size <- model$reg_table$b.r
  SE <- model$reg_table$SE
  CI_Lower <- model$reg_table$CI.L
  CI_Upper <- model$reg_table$CI.U
  P_Value <- model$reg_table$prob
  
  # Combine effect size and 95% CI into one column
  Effect_Size_CI <- paste0(
    round(Effect_Size, 2), 
    " (", 
    round(CI_Lower, 2), 
    ", ", 
    round(CI_Upper, 2), 
    ")"
  )
  
  # Extract model-level statistics
  Clusters <- model$N
  Outcomes <- model$M
  Omega_sq <- model$mod_info$omega.sq
  Tau_sq <- model$mod_info$tau.sq
  
  # Combine all results into a data frame
  results <- data.frame(
    Model = c("(Intercept - Alcohol)", "Drugs", "Marijuana", "Smoking"),  # Manually specify model names
    Effect_Size_CI = Effect_Size_CI,
    SE = round(SE, 2),
    P = formatC(P_Value, format = "fg", digits = 2),  # Scientific notation for small p-values
    Clusters = Clusters,
    Outcomes = Outcomes,
    Omega_sq = round(Omega_sq, 4),
    Tau_sq = round(Tau_sq, 4)
  )
  
  return(results)
}

# Extract the results for the meta.rob model
results_for_table <- extract_robu_results(meta.rob)

# Generate the Markdown table text
markdown_table <- kable(
  results_for_table, 
  format = "markdown", 
  align = c("l", "l", "l", "l", "l", "l", "l", "l")
)

# Add explicit newlines for proper formatting
markdown_table_text <- paste0(markdown_table, collapse = "\n")

# Print the Markdown table text to the console
cat(markdown_table_text)

```

# Sub-Studies
# Smoking
```{r}
( meta.rob.smoking = robu( yi ~ 1,
                   data = d.smoking, 
                   studynum = as.factor(authoryear),
                   var.eff.size = vi,
                   modelweights = "HIER",
                   small = F) )
print(meta.rob.smoking)
# Extract effect sizes and standard errors
effect_sizes <- d.smoking$yi
standard_errors <- sqrt(d.smoking$vi)
# Create a funnel plot
funnel(x = effect_sizes, sei = standard_errors, ylab = "Standard Error", xlab = "Effect Size")

```

# Alcohol
```{r}
result <- rma(
  yi,  # Log-transformed RRs
  vi,  # Variances of the log-transformed RRs
  data = d.alcohol,
  method = "REML"  # For random-effects model
)
summary(result)
forest(result, slab = d.alcohol$authoryear, header = T, font = 'helvetica')


( meta.rob.alcohol = robu( yi ~ 1, 
                   data = d.alcohol, 
                   studynum = as.factor(authoryear),
                   var.eff.size = vi,
                   modelweights = "HIER",
                   small = F) )
```

# Marijuana
```{r}
result <- rma(
  yi,  # Log-transformed RRs
  vi,  # Variances of the log-transformed RRs
  data = d.marijuana,
  method = "REML"  # For random-effects model
)
summary(result)
forest(result, slab = d.marijuana$authoryear, header = T, font = 'helvetica')

( meta.rob.marijuana = robu( yi ~ 1, 
                   data = d.marijuana, 
                   studynum = as.factor(authoryear),
                   var.eff.size = vi,
                   modelweights = "HIER",
                   small = FALSE) )
```

# Drugs
```{r}
result <- rma(
  yi,  # Log-transformed RRs
  vi,  # Variances of the log-transformed RRs
  data = d.drugs,
  method = "REML"  # For random-effects model
)
summary(result)
forest(result, slab = d.drugs$authoryear, header = T, font = 'helvetica')

( meta.rob.drugs = robu( yi ~ 1, 
                   data = d.drugs, 
                   studynum = as.factor(authoryear),
                   var.eff.size = vi,
                   modelweights = "HIER",
                   small = FALSE) )
```

# Substudy Markdown Table
```{r}
format_value <- function(x) {
  if (x < 0.001) {
    return("<0.001")
  } else {
    return(formatC(x, format = "fg", digits = 3))
  }
}

# Function to safely extract model results along with additional statistics
get_model_results <- function(model) {
  if (!is.null(model$reg_table)) {
    return(list(
      Effect_Size = model$reg_table$b.r,  # Extract beta from b.r
      SE = format_value(model$reg_table$SE),
      CI_Lower = model$reg_table$CI.L,
      CI_Upper = model$reg_table$CI.U,
      P_Value = format_value(model$reg_table$prob),
      Clusters = model$N,  # Number of clusters
      Outcomes = model$M,  # Number of outcomes
      Omega_sq = format_value(model$mod_info$omega.sq),  # Omega squared
      Tau_sq = format_value(model$mod_info$tau.sq)  # Tau squared
    ))
  } else {
    return(list(
      Effect_Size = NA,
      SE = NA,
      CI_Lower = NA,
      CI_Upper = NA,
      P_Value = NA,
      Clusters = NA,
      Outcomes = NA,
      Omega_sq = NA,
      Tau_sq = NA
    ))
  }
}

# Extract results safely
alcohol_results <- get_model_results(meta.rob.alcohol)
drugs_results <- get_model_results(meta.rob.drugs)
smoking_results <- get_model_results(meta.rob.smoking)
marijuana_results <- get_model_results(meta.rob.marijuana)

# Create the results data frame
results <- data.frame(
  Model = c("Alcohol", "Drugs", "Smoking", "Marijuana"),
  Effect_Size = c(alcohol_results$Effect_Size, drugs_results$Effect_Size, smoking_results$Effect_Size, marijuana_results$Effect_Size),
  SE = c(alcohol_results$SE, drugs_results$SE, smoking_results$SE, marijuana_results$SE),
  CI_Lower = c(alcohol_results$CI_Lower, drugs_results$CI_Lower, smoking_results$CI_Lower, marijuana_results$CI_Lower),
  CI_Upper = c(alcohol_results$CI_Upper, drugs_results$CI_Upper, smoking_results$CI_Upper, marijuana_results$CI_Upper),
  P_Value = c(alcohol_results$P_Value, drugs_results$P_Value, smoking_results$P_Value, marijuana_results$P_Value),
  Clusters = c(alcohol_results$Clusters, drugs_results$Clusters, smoking_results$Clusters, marijuana_results$Clusters),
  Outcomes = c(alcohol_results$Outcomes, drugs_results$Outcomes, smoking_results$Outcomes, marijuana_results$Outcomes),
  Omega_sq = c(alcohol_results$Omega_sq, drugs_results$Omega_sq, smoking_results$Omega_sq, marijuana_results$Omega_sq),
  Tau_sq = c(alcohol_results$Tau_sq, drugs_results$Tau_sq, smoking_results$Tau_sq, marijuana_results$Tau_sq)
)

# Modify the results to combine Effect Size and 95% CI into one column
results$Effect_Size_CI <- paste0(
  round(results$Effect_Size, 2), 
  " (", 
  round(results$CI_Lower, 2), 
  ", ", 
  round(results$CI_Upper, 2), 
  ")"
)

# Select the relevant columns, including the Model column
results_for_table <- results[, c("Model", "Effect_Size_CI", "SE", "P_Value", "Clusters", "Outcomes", "Omega_sq", "Tau_sq")]

# Rename columns for the table
colnames(results_for_table) <- c("Model", "Effect Size (95% CI)", "SE", "P", "Clusters", "Outcomes", "Omega.sq", "Tau.sq")

# Generate the Markdown table text
markdown_table <- kable(
  results_for_table, 
  format = "markdown", 
  align = c("l", "l", "l", "l", "l", "l", "l", "l")
)

# Add explicit newlines for proper formatting
markdown_table_text <- paste0(markdown_table, collapse = "\n")

# Print the Markdown table text to the console
cat(markdown_table_text)
```




# Exposure Category Analysis
```{r}
( meta.rob = robu( yi ~ 1 + exposure, 
                   data = d.converted, 
                   studynum = as.factor(authoryear),
                   var.eff.size = vi,
                   modelweights = "HIER",
                   small = F) )

# Extract relevant statistics
num_clusters <- meta.rob$N
num_outcomes <- meta.rob$M
min_outcomes <- min(meta.rob$studywise_stats$num_outcomes)
mean_outcomes <- round(mean(meta.rob$studywise_stats$num_outcomes), 2)
median_outcomes <- median(meta.rob$studywise_stats$num_outcomes)
max_outcomes <- max(meta.rob$studywise_stats$num_outcomes)
omega_sq <- formatC(meta.rob$mod_info$omega.sq, format = "fg", digits = 7)
tau_sq <- formatC(meta.rob$mod_info$tau.sq, format = "fg", digits = 7)

# Extract regression table results
reg_table <- meta.rob$reg_table %>%
  mutate(
    Label = c("R/S", "RSA"),
    Estimate = formatC(b.r, format = "fg", digits = 3),
    StdErr = formatC(SE, format = "fg", digits = 3),
    t_value = formatC(t, format = "fg", digits = 2),
    p_value = ifelse(prob < 0.001, "<0.001", formatC(prob, format = "e", digits = 2)),
    CI_Lower = formatC(CI.L, format = "fg", digits = 3),
    CI_Upper = formatC(CI.U, format = "fg", digits = 3)
  ) %>%
  select(Label, Estimate, StdErr, t_value, p_value, CI_Lower, CI_Upper)

# Generate a table summarizing the results
summary_table <- data.frame(
  "Number of Clusters" = num_clusters,
  "Number of Outcomes" = num_outcomes,
  "Omega.sq" = omega_sq,
  "Tau.sq" = tau_sq
)

# Print the summary table
summary_text <- kable(
  summary_table,
  format = "markdown",
  align = c("l", "r")
)
cat(summary_text, "\n")

# Generate the regression results table with labels
regression_table <- kable(
  reg_table,
  format = "markdown",
  align = c("l", "r", "r", "r", "r", "r", "r"),
  col.names = c("Label", "Estimate", "StdErr", "t-value", "P(|t|)", "95% CI.L", "95% CI.U")
)

# Add explicit newlines for proper formatting
regression_table_text <- paste0(regression_table, collapse = "\n")
regression_table_text <- gsub("\\n", "\n", regression_table_text)

# Print the regression results table with proper newlines
cat(regression_table_text, "\n")

```

```{r}
d.rsa <- d.converted %>% filter(exposure == 'RSA')
( meta.rob = robu( yi ~ 1, 
                   data = d.rsa, 
                   studynum = as.factor(authoryear),
                   var.eff.size = vi,
                   modelweights = "HIER",
                   small = F) )

d.rs  <- d.converted %>% filter(exposure == 'R/S')
( meta.rob = robu( yi ~ 1, 
                   data = d.rs, 
                   studynum = as.factor(authoryear),
                   var.eff.size = vi,
                   modelweights = "HIER",
                   small = F) )


```

# Bias
```{r}
d.all$bias2<- d.all$bias
d.all$bias2[ grepl('Low',      d.all$bias) ]<- 'Low'
d.all$bias2[ grepl('Moderate', d.all$bias) ]<- 'Moderate'
d.all$bias2[ grepl('Serious',  d.all$bias) ]<- 'Serious'

( meta.rob.bias = robu( yi ~ 1 + bias2, 
                   data = d.all, 
                   studynum = as.factor(authoryear),
                   var.eff.size = vi,
                   modelweights = "HIER",
                   small = F) )

# Vectorized helper function to format numbers with 3 significant digits
# If any value < 0.001, return "<0.001"
format_value <- function(x) {
  ifelse(x < 0.001, "<0.001", formatC(x, format = "fg", digits = 3))
}

# Function to extract and format the robu model results
extract_robu_results_bias <- function(model) {
  # Extract the effect size, SE, CI, and p-value
  Effect_Size <- model$reg_table$b.r
  SE <- model$reg_table$SE
  CI_Lower <- model$reg_table$CI.L
  CI_Upper <- model$reg_table$CI.U
  P_Value <- model$reg_table$prob
  
  # Combine effect size and 95% CI into one column with 3 significant digits
  Effect_Size_CI <- paste0(
    round(Effect_Size, 2), 
    " (", 
    round(CI_Lower, 2), 
    ", ", 
    round(CI_Upper, 2), 
    ")"
  )
  
  # Extract model-level statistics
  Clusters <- model$N
  Outcomes <- model$M
  Omega_sq <- model$mod_info$omega.sq
  Tau_sq <- model$mod_info$tau.sq
  
  # Combine all results into a data frame
  results <- data.frame(
    Model = c("(Intercept - Low Bias)", "Moderate Bias", "Serious Bias"),  # Manually specify model names
    Effect_Size_CI = Effect_Size_CI,
    SE = format_value(SE),
    P = format_value(P_Value),
    Clusters = Clusters,
    Outcomes = Outcomes,
    Omega_sq = format_value(Omega_sq),
    Tau_sq = format_value(Tau_sq)
  )
  
  return(results)
}

# Extract the results for the meta.rob model
results_for_table_bias <- extract_robu_results_bias(meta.rob.bias)

# Generate the Markdown table text
markdown_table_bias <- kable(
  results_for_table_bias, 
  format = "markdown", 
  align = c("l", "l", "l", "l", "l", "l", "l", "l")
)

# Add explicit newlines for proper formatting
markdown_table_bias_text <- paste0(markdown_table_bias, collapse = "\n")

# Print the Markdown table text to the console
cat(markdown_table_bias_text)
# Optionally, copy to clipboard using pbcopy (for macOS)
cat(egger_test_filtered, file = pipe("pbcopy"))
```

# Headline Analysis
```{r}
d.all$headline2 <- grepl('yes', d.all$headline)
( meta.rob.headline = robu( yi ~ 1 + headline2, 
                   data = d.all, 
                   studynum = as.factor(authoryear),
                   var.eff.size = vi,
                   modelweights = "HIER",
                   small = F) )

# Vectorized helper function to format numbers with 3 significant digits
# If any value < 0.001, return "<0.001"
format_value <- function(x) {
  ifelse(x < 0.001, "<0.001", formatC(x, format = "fg", digits = 3))
}

# Function to extract and format the robu model results
extract_robu_results_headline <- function(model) {
  # Extract the effect size, SE, CI, and p-value
  Effect_Size <- model$reg_table$b.r
  SE <- model$reg_table$SE
  CI_Lower <- model$reg_table$CI.L
  CI_Upper <- model$reg_table$CI.U
  P_Value <- model$reg_table$prob
  
  # Combine effect size and 95% CI into one column with 3 significant digits
  Effect_Size_CI <- paste0(
    round(Effect_Size, 2), 
    " (", 
    round(CI_Lower, 2), 
    ", ", 
    round(CI_Upper, 2), 
    ")"
  )
  
  # Extract model-level statistics
  Clusters <- model$N
  Outcomes <- model$M
  Omega_sq <- model$mod_info$omega.sq
  Tau_sq <- model$mod_info$tau.sq
  
  # Combine all results into a data frame
  results <- data.frame(
    Model = c("(Intercept - No Headline)", "Headline"),  # Manually specify model names
    Effect_Size_CI = Effect_Size_CI,
    SE = format_value(SE),
    P = format_value(P_Value),
    Clusters = Clusters,
    Outcomes = Outcomes,
    Omega_sq = format_value(Omega_sq),
    Tau_sq = format_value(Tau_sq)
  )
  
  return(results)
}

# Extract the results for the meta.rob.headline model
results_for_table_headline <- extract_robu_results_headline(meta.rob.headline)

# Generate the Markdown table text
markdown_table_headline <- kable(
  results_for_table_headline, 
  format = "markdown", 
  align = c("l", "l", "l", "l", "l", "l", "l", "l")
)

# Add explicit newlines for proper formatting
markdown_table_headline_text <- paste0(markdown_table_headline, collapse = "\n")

# Print the Markdown table text to the console
cat(markdown_table_headline_text)

# Optionally, copy to clipboard using pbcopy (for macOS)
cat(markdown_table_headline_text, file = pipe("pbcopy"))


```
# Bootstrap
```{r}
run_rma_on_subset <- function(data, x_percent) {
  subset_data <- data[sample(1:nrow(data), size = floor(x_percent * nrow(data))), ]
  model <- rma(yi, vi, data = subset_data) # Adjust the rma function as needed
  return(coef(model))
}

run_robu_on_subset <- function(data, x_percent) {
  subset_data <- data[sample(1:nrow(data), size = floor(x_percent * nrow(data))), ]
  #model <- rma(yi, vi, data = subset_data) # Adjust the rma function as needed
  #return(coef(model))
  meta.rob = robu( yi ~ 1, 
                   data = subset_data,
                   studynum = as.factor(authoryear),
                   var.eff.size = vi,
                   modelweights = "HIER",
                   small = F) 
  return(meta.rob$b.r)
}

# Sensitivity analysis parameters
set.seed(123) # For reproducibility
x_percentages <- c(0.5, 0.6, 0.7, 0.8, 0.9, 1.0)
n_iterations <- 1000
results_rma <- list()
results_robu <- list()

# Run analysis for rma
for (x in x_percentages) {
  estimates <- replicate(n_iterations, run_rma_on_subset(d.converted, x))
  results_rma[[as.character(x)]] <- estimates
}

# Run analysis for robu
for (x in x_percentages) {
  estimates <- replicate(n_iterations, run_robu_on_subset(d.converted, x))
  results_robu[[as.character(x)]] <- estimates
}

# Calculate averages and CIs for rma
avg_estimates_rma <- sapply(results_rma, mean)
ci_lower_rma <- sapply(results_rma, function(x) quantile(x, 0.025))
ci_upper_rma <- sapply(results_rma, function(x) quantile(x, 0.975))

# Calculate averages and CIs for robu
avg_estimates_robu <- sapply(results_robu, mean)
ci_lower_robu <- sapply(results_robu, function(x) quantile(x, 0.025))
ci_upper_robu <- sapply(results_robu, function(x) quantile(x, 0.975))

```

# Plot Bootstrap Results
```{r}
# Set up the file to save the image
png("sensitivity_analysis_two_panel.png", width = 10, height = 5, units = "in", res = 300)  # 10 inches wide, 5 inches tall, 300 DPI

# Create a two-panel figure
par(mfrow = c(1, 2))  # Set up a 1x2 plotting area

# Small offset to use for zero-length arrows
offset <- 0.001

# Define y-axis limits
y_limits <- c(-0.25, 0)

# Panel A: Plot for rma
plot(x_percentages, avg_estimates_rma, ylim=y_limits, type="b",
     xlab="Percentage of Data", ylab="Effect Size Estimate", main="Panel A: rma")
for (i in 1:length(x_percentages)) {
  lower <- ci_lower_rma[i]
  upper <- ci_upper_rma[i]
  
  # Draw arrows if CI is not zero-length
  if (lower != upper) {
    arrows(x_percentages[i], lower, x_percentages[i], upper, angle=90, code=3, length=0.1)
  } else {
    # Draw a small horizontal line if CI is zero-length
    segments(x_percentages[i] - offset, lower, x_percentages[i] + offset, upper)
  }
}

# Panel B: Plot for robu
plot(x_percentages, avg_estimates_robu, ylim=y_limits, type="b",
     xlab="Percentage of Data", ylab="Effect Size Estimate", main="Panel B: robu")
for (i in 1:length(x_percentages)) {
  lower <- ci_lower_robu[i]
  upper <- ci_upper_robu[i]
  
  # Draw arrows if CI is not zero-length
  if (lower != upper) {
    arrows(x_percentages[i], lower, x_percentages[i], upper, angle=90, code=3, length=0.1)
  } else {
    # Draw a small horizontal line if CI is zero-length
    segments(x_percentages[i] - offset, lower, x_percentages[i] + offset, upper)
  }
}

# Close the file to finalize the plot
dev.off()


```

# Worst Case Meta-analysis
```{r}
# Subset for non-significant results or effects in the opposite direction

# TODO: go back and add p-values during the etl. For now, doing a simple loop
d.all$pval<-NA
for(i in 1:dim(d.all)[1]){
  ay <- d.all$id[i]
  found<- df.substance %>% filter( id == ay)
  p<- found$PValue[1]
  numericPValue <- as.numeric(gsub("[^0-9.]", "", p))
  if(substr(p[1], 1, 1) == '>'){
    print('found a >')
    numericPValue<- numericPValue+0.01
    print(numericPValue)
  }
  d.all$pval[i]<- numericPValue
}


#d.all$pval <- cleanPValue(d.all$pval)

d.worst_case3<- d.all %>% filter(yi > 0 | pval > 0.05)


( meta.rob = robu( yi ~ 1, 
                   data = d.worst_case3,
                   studynum = as.factor(authoryear),
                   var.eff.size = vi,
                   modelweights = "HIER",
                   small = F) )


```

# Hedge's Selection
```{r}
# Compute z-scores
z_scores <- d.all$yi / sqrt(d.all$vi)

# Compute one-tailed p-values, which is what weightfunct uses
d.all$pval_one_tailed <- ifelse(
  d.all$yi >= 0,
  1 - pnorm(z_scores),
  pnorm(z_scores)
)
hist(d.all$pval_one_tailed, nbins=20)

# Define p-value intervals (use fewer, broader intervals to avoid empty categories)
p_value_intervals <- c(0.025, 0.10)

# Check how many p-values fall into the interval
table(cut(d.all$pval, breaks = p_value_intervals, include.lowest = TRUE))

# Define weights for sensitivity analysis
specified_weights <- c(10, 0.5, 0.5)

# Run the Vevea and Hedges weight-function model
# Specifiying pvalues and steps results in NaNs
hedges_model <- weightfunct(
  effect = d.all$yi,
  v = d.all$vi,
  steps = p_value_intervals,
  pval = d.all$pval_one_tailed,
  table = TRUE  # Print table of p-value intervals and effect sizes per interval
)
(hedges_model)

# PLAYING AROUND WITH DEFINING WEIGHTS
p_value_intervals <- c(0.025, 0.10)
specified_weights <- c(10, 0.5, 0.5)
hedges_model <- weightfunct(
  effect = d.all$yi,
  v = d.all$vi,
  steps = p_value_intervals,
  pval = d.all$pval_one_tailed,
  weights = specified_weights,
  table = TRUE  # Print table of p-value intervals and effect sizes per interval
)
(hedges_model)

# From Maya
temp = rma( yi = d.all$yi,
           vi = d.all$vi,
           method = "REML",
           knha = TRUE )
metafor::selmodel(temp, type="stepfun", alternative = "greater", steps=c(0.025, 1))

( m1 = weightfunct( effect = d.all$yi,
                    v = d.all$vi,
                    steps = c(0.025, 1),
                    table = TRUE ) )
(H = m1[[2]]$hessian)
(ses = sqrt( diag( solve(H) ) ))
(lrchisq <- 2 * (abs(m1[[1]]$value - m1[[2]]$value)))
(LR_pval <- 1 - pchisq(lrchisq, 1L))
# selection ratio
# index [3] would change for PSM with more cut points
1/m1$output_adj$par[3]

```
# MetaUtility
```{r}
calibrated<- calib_ests(d.all$yi, sqrt(d.all$vi), method = "DL")

plot(calibrated)


```

# "Best"
```{r}
d.best <- d.all %>% filter(bias2 != 'Serious')
d.best <- d.best %>% filter(bias2 != 'Critical')
d.best <- d.best %>% filter(exposure == 'RSA')

( meta.rob = robu( yi ~ 1, 
                   data = d.best,
                   studynum = as.factor(authoryear),
                   var.eff.size = vi,
                   modelweights = "HIER",
                   small = F) )

```
